Making ice sheet models \\ scale properly

### 2023-01-05 talk/slides.tex
progress finally

\usepackage{empheq,bm}
\usepackage{xspace}
\usepackage{fancyvrb}

@@

\begin{columns}
\begin{column}{0.6\textwidth}
\begin{itemize}
\item for simplicity/clarity of the upcoming performance analysis, I will be ignoring quite a bit of physics!
\item \alert{ignoring}:
    \begin{itemize}
    \item[$\circ$] floating ice
    \item[$\circ$] subglacial hydrology
    \item[$\circ$] ice temperature
    \item[$\circ$] fracture processes (calving, crevasses)
    \end{itemize}
\item now, a simple definition of ``ice sheet model'' \dots
\end{itemize}
\end{column}
\begin{column}{0.42\textwidth}
\hfill \includegraphics[width=0.9\textwidth]{images/map-glacier-ice.png}
\end{column}
\end{columns}
\end{frame}

@@

\begin{columns}
\begin{column}{0.6\textwidth}
\begin{itemize}
\begin{definition}
an \alert{ice sheet model} is a \underline{map} which simulates an ice sheet in a climate
\end{definition} 
\item two inputs:
    \begin{itemize}
    \item[$\circ$] \emph{surface mass balance}
$$a(t,x) = \begin{pmatrix}
\text{balance of solid precipitation} \\
\text{\, vs melt \& runoff at surfaces}
\end{pmatrix}$$

@@

    \item[$\circ$] \emph{bed elevation} $b(x)$
    \end{itemize}
\item two outputs:
    \begin{itemize}
    \item[$\circ$] \emph{upper surface elevation} $s(t,x)$
    \item[$\circ$] \emph{ice velocity} $\bu(t,x)$
    \end{itemize}
\item {\small $\begin{pmatrix} \text{climate} \\ \text{topography} \end{pmatrix} \to \begin{pmatrix} \text{geometry} \\ \text{velocity} \end{pmatrix}$}
\end{itemize}
\end{column}
\begin{column}{0.42\textwidth}
\hfill \includegraphics[width=0.9\textwidth]{images/map-velocity.png}

@@

@@

\begin{itemize}
\item as this is a mathematics talk, more or less, some notation \dots
\item FIXME conservation laws; conserve momentum standard; conserve mass means both conserve mass in interior (incompressible) and at surface (SKE)
\end{itemize}
\end{frame}

>>>>

\usepackage{empheq,bm,xspace,fancyvrb,soul}

@@

\begin{columns}
\begin{column}{0.6\textwidth}
\begin{itemize}
\item for simplicity/clarity of the upcoming performance analysis, I will be ignoring quite a bit of physics!
\item \alert{ignoring}:
    \begin{itemize}
    \item[$\circ$] floating ice
    \item[$\circ$] subglacial hydrology
    \item[$\circ$] ice temperature
    \item[$\circ$] fracture processes (calving, crevasses)
    \item[$\circ$] solid earth deformation
    \end{itemize}
\item now, a simple definition of ``ice sheet model'' \dots
\end{itemize}
\end{column}
\begin{column}{0.42\textwidth}
\hfill \includegraphics[width=0.9\textwidth]{images/map-glacier-ice.png}
\end{column}
\end{columns}
\end{frame}

@@

\begin{columns}
\begin{column}{0.6\textwidth}
\begin{itemize}
\begin{definition}
an \alert{ice sheet model} is a \underline{map} which simulates an ice sheet in a climate
\end{definition} 
\item two inputs:
    \begin{itemize}
    \item[$\circ$] \emph{surface mass balance}
$$\hspace{-7mm} a(t,x,y) = \begin{pmatrix}
\text{balance of precipitation vs} \\
\text{melt \& runoff from surfaces}
\end{pmatrix}$$

@@

    \item[$\circ$] \emph{bed elevation} $b(x,y)$
    \end{itemize}
\item two outputs:
    \begin{itemize}
    \item[$\circ$] \emph{upper surface elevation} $s(t,x,y)$
    \item[$\circ$] \emph{ice velocity} $\bu(t,x,y,z)$
    \end{itemize}
\item {\small $\begin{pmatrix} \text{climate} \\ \text{topography} \end{pmatrix} \to \begin{pmatrix} \text{geometry} \\ \text{velocity} \end{pmatrix}$}
\end{itemize}
\end{column}
\begin{column}{0.42\textwidth}
\hfill \includegraphics[width=0.9\textwidth]{images/map-velocity.png}

@@

\begin{frame}{notation}

\begin{itemize}
%\item but this is a mathematics talk \dots notation! equations!
\item the data $a(t,x,y)$, $b(x,y)$ is defined on fixed time and space domains:
	$$t \in [0,T] \quad \text{and} \quad (x,y) \in \Omega \subset \RR^2$$

\begin{center}
\includegraphics<1>[width=0.55\textheight]{images/domain-data.png}
\includegraphics<2>[width=0.6\textheight]{images/domain-velocity.png}
\end{center}

\medskip
\item<2> the solution surface elevation $s(t,x,y)$ is defined on $[0,T]\times \Omega$
    \begin{itemize}
    \item[$\circ$] $s=b$ where there is no ice
    \end{itemize}

\item<2> one must \emph{solve} for the time-dependent \alert{icy domain} $\Lambda(t) \subset \Omega$ on which the velocity $\bu(t,x,y,z)$ is defined:
    $$\Lambda(t) = \left\{(x,y,z)\,:\,b(x,y) < z < s(t,x,y)\right\}$$

\vspace{-2mm}
    \begin{itemize}
    \item[$\circ$] $\Gamma_s(t),\Gamma_b(t) \subset \partial \Lambda(t)$ denote the surface and base of the ice
    \end{itemize}
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item ice sheet evolution is a \alert{free-boundary conservation} problem
    \begin{itemize}
    \item[$\circ$] mass conservation
    \item[$\circ$] momentum conservation
    \item[$\circ$] energy conservation (\emph{ignored here})
    \end{itemize}
\item key idea: conservation of mass means both conserve mass in interior (incompressible) and at surface (SKE)
\end{itemize}
\end{frame}
